Linear regression for approximating real and synthetic data

### 2022-10-10 doc/paper.tex
More paper stuff

@@

Using Ridge regression, we see from Fig.~\ref{frankiedifflambdas} that larger and larger values of $\lambda$ efficiently prevents overfitting and reduces the variance between predictions. Looking at the beta progression with for the same values of $\lambda$ (Fig.~\ref{frankeridgebetas}), we see that the regularization parameter is successfully shrinking the beta values. Using Scikit's gridsearch, we find the optimal $\lambda$ to be $10^{-8}$ and the optimal polynomial degree to be 7, with an average MSE of $0.0037999$, which is very slightly lower than OLS for the Franke function with an added stochastic noise $\epsilon \sim N(0,0.05)$. We see from this that Ridge regression has not really had much of an effect in decreasing the MSE. We must now remember what Ridge aims to do. With the regularization parameter, we can prevent overfitting if our model is too complex for our data. But we should recall that we decide the complexity of our model, by deciding which polynomial degree we use when creating the design matrix. In fact, by choosing polynomial degree 7 for our OLS method to get the best MSE for noisy data, we are in fact doing a primitive form of best subset selection, as described in Hastie et Al. p 57 \cite{hastie}. Almost by definition, the polynomial degree with the best MSE is not overfitted, because if it was, it would not be the model with the lowest MSE among all polynomial degrees. Therefore it is not so unexpected that Ridge regression has a relatively low impact on the MSE. We could expect that Ridge would allow us to reach higher polynomial degrees without overfitting, by shrinking features of the higher polynomial degrees that are not supported but allowing us to keep other high degree terms (for example, allowing us to use $x^5y^4$, but discard $x^9$, if only $x^5y^4$ was supported by the data). But we can see from the Table. ~\ref{ols_vs_ridge_table} that we are not able to push into higher degrees. A possible explanation for this is that every feature is useful, due to the polynomial expansion of the Franke function.

@@

\bibliographystyle{alpha}
\bibliography{bibliography}

>>>>

The correctness of our linear regression can be confirmed by running our debug function (the so called SkrankeFunction)\footnote{this can be done by running \texttt{\$ python3 task\_b.py --debug -n 2}}, which is a second degree polynomial where $b_0 = 0, \; b_1 = 1, \; b_2 = 2$ and so on.

@@

Using Ridge regression, we see from Fig.~\ref{frankiedifflambdas} that larger and larger values of $\lambda$ efficiently prevents overfitting and reduces the variance between predictions. Looking at the beta progression with for the same values of $\lambda$ (Fig.~\ref{frankeridgebetas}), we see that the regularization parameter is successfully shrinking the beta values. Using Scikit's gridsearch, we find the optimal $\lambda$ to be $10^{-8}$ and the optimal polynomial degree to be 7, with an average MSE of $0.0037999$, which is very slightly lower than OLS for the Franke function with an added stochastic noise $\epsilon \sim N(0,0.05)$. We see from this that Ridge regression has not really had much of an effect in decreasing the MSE. We must now remember what Ridge aims to do. With the regularization parameter, we can prevent overfitting if our model is too complex for our data. But we should recall that we decide the complexity of our model, by deciding which polynomial degree we use when creating the design matrix. In fact, by choosing polynomial degree 7 for our OLS method to get the best MSE for noisy data, we are in fact doing a primitive form of best subset selection, as described in Hastie et Al.\cite[57]{hastie}. Almost by definition, the polynomial degree with the best MSE is not overfitted, because if it was, it would not be the model with the lowest MSE among all polynomial degrees. Therefore it is not so unexpected that Ridge regression has a relatively low impact on the MSE. We could expect that Ridge would allow us to reach higher polynomial degrees without overfitting, by shrinking features of the higher polynomial degrees that are not supported but allowing us to keep other high degree terms (for example, allowing us to use $x^5y^4$, but discard $x^9$, if only $x^5y^4$ was supported by the data). But we can see from the Table. ~\ref{ols_vs_ridge_table} that we are not able to push into higher degrees. A possible explanation for this is that every feature is useful, due to the polynomial expansion of the Franke function.

@@

\bibliographystyle{apalike}
\bibliography{bibliography}
